NLIZE: A Perturbation-Driven Visual Interrogation Tool for Analyzing and Interpreting Natural Language Inference Models

### 2018-03-30 paper/application.tex
finish section 6.1

\section{Application Scenarios}
\label{sec:caseStudy}
% \shusen{A full case study that driven by the visualization task and the question associated with them}
%In this section, we discuss application scenarios, in which the domain experts utilize the
To better illustrate how the proposed perturbation-driven exploration tool helps researchers interpret the neural network model, we present five application scenarios gathered by the domain experts who integrated the proposed tool into their analysis workflow.

\subsection{Scenario 1: Assess the Model Prediction Stability}
The robustness of the prediction is often hard to evaluate, however, they provide valuable information for the researchers to better understand the model.
%
In the proposed work, we approach the prediction robustness from sensitivity analysis point of view. The stability of the prediction is measured by how often the predicted labels are altered after small perturbations are applied to the input.
%
Compare to other types of input (e.g., image), perturbation of the natural language can be particular tricky, as small alteration of words can drastically change the meaning of the sentence. As discussed in Section~\ref{sec:sentence}, we try to maintain semantic of the sentence by only replace words with their synonymous and only replace one word for each pair.
As illustrated in Fig.~\ref{fig:predictStability}, by utilizing the proposed tool, the domain expert can not only examine visual summary of the stability, but also quickly dive into individual examples for case by case analysis.

\begin{figure}[htbp]
\centering
\vspace{-2mm}
 \includegraphics[width=1.0\linewidth]{predictStability}
 \vspace{-4mm}
 \caption{
Prediction stability assessment. In (a)(b)(c), we estimate the overall prediction stability (regarding synonymous perturbation) for each type prediction over the entire development set (10k examples). The user can drill down to individual examples by filtering via the histogram or scatterplot (see Fig.~\ref{fig:summaryView}). For highly unstable prediction, we often observe the original prediction is near a decision boundary (e.g., the yellow circle in (e) corresponds to a prediction that is at the boundary between \emph{entailment} and \emph{neutral}), however, some predictions, such as the one illustrated in (d) can alter the prediction quite drastically with minor perturbation (illustrated in $d_1$, where the word ``pile'' is replace by ``heap'' in the hypothesis sentence).
}
\label{fig:predictStability}
\vspace{-2mm}
\end{figure}

In Fig.~\ref{fig:predictStability}(a)(b)(c),  %by viewing the distribution of the stability in the histogram
we compare the overall prediction stability (regarding synonymous perturbation) for each type prediction over the entire development set (10k examples).
%
We observe a drastic difference between the stability for entailment predictions compare to the contradiction and neutral ones.
%
Such a distinction can be partially explained by how entailment relationship is defined. As discussed in the Section~\ref{sec:languageInference}, the relationship is only valid if the concept in premise is more specific than the concept in hypothesis (therefore, you can infer the hypothesis from the premise, but not other way around). This means, we may change the entailment relationships simply by replacing nouns and verb by synonymous, whereas, the same does not apply for neutral and contradiction.
%
The inherent sensitivity difference may motivate domain experts designing dedicate model mechanism to address the such a disparity.

Beside presenting the summary view, the tool also allow user to quickly narrow down the select to single example  by filtering through the histogram and scatterplot (see Fig.~\ref{fig:summaryView}).
%
Through exploring multiple samples (in the entailment category, E/E in Fig.~\ref{fig:summaryView}(a)), the domain experts noticed many highly unstable outliers are from sentence pairs where the prediction is near the decision boundary (see Fig.~\ref{fig:predictStability}(e), the yellow circle corresponds to a \emph{entailment} prediction that is very close to \emph{neutral}).
However, some predictions, such as the one illustrated in Fig.~ref{fig:predictStability}(d) can alter the prediction quite drastically with minor perturbation (illustrated in $d_1$, where the word ``heap'' is replace by ``pile'' in the hypothesis sentence. We will try to understand the what is happened inside the model in the following section (see Fig.~\ref{fig:att2pred}).

@@

\begin{figure}[htbp]
\centering
\vspace{-2mm}
 \includegraphics[width=1.0\linewidth]{wrongAlignRightPred}
 \caption{
The example of model producing right prediction with wrong attention.
The word \textbf{break} and \textbf{sit} should not align with each other (a), yet, it still yield the correct prediction (b).
}
\label{fig:wrongAlignRightPred}
\end{figure}

>>>>

\section{Application Scenarios}
\label{sec:caseStudy}
% \shusen{A full case study that driven by the visualization task and the question associated with them}
%In this section, we discuss application scenarios, in which the domain experts utilize the
To better illustrate how the proposed perturbation-driven exploration tool helps researchers interpret the neural network model, we present five application scenarios gathered by the domain experts who integrated the proposed tool in their analysis workflow.

\subsection{Scenario 1: Assess the Model Prediction Stability}
The robustness of the prediction is often hard to evaluate. However, the prediction stability provides valuable information for the researchers to better understand the model.
%
In the proposed work, we approach the prediction robustness from sensitivity analysis point of view. The stability of the prediction is measured by how often the predicted labels are altered after small perturbations are applied to the input.
%
Compare to other types of input (e.g., image), perturbation of the natural language can be particularly tricky, as small alteration of words can drastically change the meaning of the sentence. As discussed in Section~\ref{sec:sentence}, we try to maintain semantic of the sentence by only replace words with their synonymous and only replace one word for each pair.
As illustrated in Fig.~\ref{fig:predictStability}, by utilizing the proposed tool, the domain expert can not only examine a visual summary of the stability but also quickly dive into individual examples for case by case analysis.

\begin{figure}[htbp]
\centering
\vspace{-2mm}
 \includegraphics[width=1.0\linewidth]{predictStability}
 \vspace{-4mm}
 \caption{
Prediction stability assessment. In (a)(b)(c), we estimate the overall prediction stability (regarding synonymous perturbation) for each type prediction over the entire development set (10k examples). The user can drill down to individual examples by filtering via the histogram or scatterplot (see Fig.~\ref{fig:summaryView}). For highly unstable prediction, we often observe the original prediction is near a decision boundary (e.g., the yellow circle in (e) corresponds to a prediction that is at the boundary between \emph{entailment} and \emph{neutral}), however, some predictions, such as the one illustrated in (d) can alter the prediction quite drastically with minor perturbation (illustrated in $d_1$, where the word \textbf{pile} is replaced by \textbf{heap} in the hypothesis sentence).
}
\label{fig:predictStability}
\vspace{-2mm}
\end{figure}

In Fig.~\ref{fig:predictStability}(a)(b)(c), we compare the overall prediction stability (regarding synonymous perturbation) for all correct predictions in the development set (10k examples in total).
%
We observe a drastic difference between the stability for \emph{entailment} predictions compare to the \emph{contradiction} and \emph{neutral} ones.
%
Such a distinction can be partially explained by how \emph{entailment} relationship is defined. The relationship is only valid if the concept in the premise is more specific than the concept in the hypothesis. Therefore, the synonymous perturbation may change the \emph{entailment} relationship, as the replaced noun and verb can be more or less restrictive compared to the original.
This inherent disparity of sensitivity may worth be taken into consideration when designing future NLI models.

Besides presenting the summary view, the tool also allows the user to quickly narrow down the selection to single example by filtering via the histogram and scatterplot (see Fig.~\ref{fig:summaryView}).
%
Through exploration of many samples with low stabilities (i.e., high sensitivities values in the plots), the domain experts noticed that many highly unstable outliers are from sentence pairs where the prediction is near the decision boundary (see Fig.~\ref{fig:predictStability}(e), the yellow circle corresponds to a \emph{entailment} prediction that is very close to \emph{neutral}).
However, some predictions, such as the one illustrated in Fig.~\ref{fig:predictStability}(d) can alter the prediction quite drastically with minor perturbation (shown in $d_1$, where the word \textbf{heap} is replaced by \textbf{pile} in the hypothesis sentence. In the following section, we will examine what happened inside the model and hypothesis on the cause of failure (see Fig.~\ref{fig:att2pred}).

@@

\begin{figure}[htbp]
\centering
\vspace{-2mm}
 \includegraphics[width=1.0\linewidth]{wrongAlignRightPred}
 \caption{
The model may produces the right prediction for the ``wrong'' reason, i.e., attention.
The word \textbf{break} and \textbf{sit} should not align with each other (a). Yet, based on the attention, model still produces the correct prediction (b).
}
\label{fig:wrongAlignRightPred}
\end{figure}
